Optimal Algorithms for Dynamic Balanced Graph Partitioning

### 2020-05-14 paper/disc20clustering.tex
Unfinished phase

	We proceed in similar steps by joining the current pivot node to a ground set of the same origin residing in a different cluster.
	Consider the step $i$,
	when the isolated node $x_i$ is collocated with $B$.
	 We issue a request between $x_i$ and some node in $C_i$,
	 where $C_i$ is the largest ground set
	 s.t.~$I(C_i) = I(x_i), C_i \neq \{x_0,y_0\}$.
%	 only if  a feasible partition exists
%	 that collocates all nodes in $\{x_i\} \cup C_i $.
	Any feasible partition replaces $x_i$ with some isolated node $x_{i+1}$,
	as the new ground set $\{x_i\} \cup C_i $  may not be ever split.
%	Otherwise,
%	no such feasible partition exists and
%	we terminate the process.
	We terminate the process once the number of remaining isolated nodes 
	is less than $\ell+3$.
	At each step $i$,
	the number of isolated nodes decrease by one or two.
	Therefore, once the process terminates,
	there are at least $\ell+1$ isolated nodes left.

@@

	\ALG performs at least one swap at each step $i$,
	and some ground set grows.
%	In other words,
%	some ground set grows by one node at each of these steps.
	Consider any ground set $C^* \neq B$ after the termination.
	This ground set has grown exactly $|C^*|-1$ times until the the termination. 
	Let $\mathcal{S}$ be the set of all ground sets after the process terminates.
	Thus, $\mathcal{S}$ includes ground sets $B$,
	$\{x_0,y_0\}$,
	and (up to) $\ell+2$ singleton ground sets.
	Among the remaining ground sets in $\mathcal{S}$,
	no two ground sets have the same origin.
	Otherwise,
   the smaller ground set is either a singleton,
   which contradicts the bound $\ell+2$ on the number of singletons,
	or we have joined nodes to it at some step,
	contradicting our choice of the largest $C_i$ at step $i$.
	Hence,  there are at most $\ell-1$ such grown ground sets,
	one per cluster,
	 excluding the cluster containing $B$. 
	Therefore,  $|\mathcal{S}| \leq 1 + 1 + (\ell+2)  + (\ell-1) = 2\ell+3$.
	Note that among all non-singleton ground sets in $\mathcal{S}$,
	 only $B$ does not grow during the process.
	 Thus,
    the total number of times that 
	a ground set in $\mathcal{S}$  has grown is
	$	\sum_{C^* \in \mathcal{S} } (|C^*|-1) - (k-1) $
	\begin{align*}
	=\sum_{C^* \in \mathcal{S} } |C^*| - 
		\sum_{C^* \in \mathcal{S} } 1 - (k-1) 
	\geq  k  \ell - (2\ell+3) - (k-1) =  (k-2)(\ell-1) - 4,
	\end{align*}
	which bounds the number of swaps performed by \ALG.
	The competitive ratio is then
	$\ALG / \OPT \geq ((k-2)(\ell-1) - 4) / 2$.
	%%
	\begin{comment}
	\ALG performs at least one swap  every time   a~ground set is constructed.
	Each expansion reduces the number of isolated nodes by $1$ or $2$.
	The decrease of two isolated nodes occurs when the largest component on  $I(x_i)$ is still singleton; this occurs at most $\ell$ times.
	We start with $k \cdot \ell - (k-1)$ isolated nodes (a~single $k-1$ ground set is revealed prior to $\{x_0, y_0\}$), and we end with isolated $\ell+1$ nodes.
	Hence, the total number of swaps of \ALG is at least $k\cdot \ell - k - 2 \ell$.
	The competitive ratio is then $\ALG / \OPT \geq (k\cdot \ell - k - 2\ell) / 2$.
	\end{comment}
	%%
\end{proof}

@@

	Finally, to bound the competitive ratio, we transform the above inequality in the following way: $|\sigma_E| \leq 2(\OPT(\sigma_E)+\OPT(M)) + 4\alpha \cdot \ell$.
	For succintness, let $\xi := \OPT(\sigma_E) + \OPT(M)$.
	Combining the bounds on the cost of \TAlg and \OPT during each finished phase, the competitive ratio is
%
	\[
		\frac{\TAlg(\sigma)}{\OPT(\sigma)} \leq \frac{11\alpha \cdot \ell + |\sigma_E|}{\alpha/2 + \xi/2} \leq \frac{30\alpha\cdot\ell + 4\cdot \xi}{\alpha + \xi} \leq 30 \ell.
	\]
%	
	\medskip

	It remains to consider the last, unfinished phase.
	First, consider the case, where the unfinished phase is also the first one.
	Then, we cannot charge $\OPT$ due to the inability to partition the components.
	Instead, we use the fact that \TAlg and $\OPT$ started with the same initial partition.
	If the input finished before the first $\alpha$ external requests, then \TAlg is $1$-competitive.
	If at least $\alpha$ external requests were issued, then $\OPT$ either paid $\alpha$ for serving them remotely, or paid $\alpha$ for a~migration; for the cost of \TAlg we follow the analysis of the unsaturated pairs.
	Second, consider the case, where there are at least two phases, then we split the cost $\alpha$ charged in the penultimate phase into last two phases, and follow the analysis regarding the unsaturated requests.
	This way, the competitive ratio increases at most twofold in comparison to a~finished phase, and the competitive ratio is $\TAlg(\sigma)/\OPT(\sigma) \leq 60\ell$.
\end{proof}

>>>>

	We proceed in similar steps by joining the current pivot node to a ground set of the same origin residing in a different cluster.
	Consider the step $i$,
	when the isolated node $x_i$ is collocated with $B$.
	 We issue a request between $x_i$ and some node in $C_i$,
	 where $C_i$ is the largest ground set
	 s.t.~$I(C_i) = I(x_i), C_i \neq \{x_0,y_0\}$ \maciek{Ground sets merges instead of requests}.
%	 only if  a feasible partition exists
%	 that collocates all nodes in $\{x_i\} \cup C_i $.
	Any feasible partition replaces $x_i$ with some isolated node $x_{i+1}$,
	as the new ground set $\{x_i\} \cup C_i $  may not be ever split.
%	Otherwise,
%	no such feasible partition exists and
%	we terminate the process.
	We terminate the process once the number of remaining isolated nodes 
	is less than $\ell+3$.
	At each step $i$,
	the number of isolated nodes decrease by one or two \maciek{Is this clear why? I am not, sure, take a look and decide. Possibly we need to say about two singletons merge}.
	Therefore, once the process terminates,
	there are at least $\ell+1$ isolated nodes left.

@@

	\ALG performs at least one swap at each step $i$,
	and some ground set grows.
%	In other words,
%	some ground set grows by one node at each of these steps.
	Consider any ground set $C^* \neq B$ after the termination.
	This ground set has grown exactly $|C^*|-1$ times until the the termination. 
	Let $\mathcal{S}$ be the set of all ground sets after the process terminates.
	Thus, $\mathcal{S}$ includes ground sets $B$,
	$\{x_0,y_0\}$,
	and (up to) $\ell+2$ singleton ground sets.
	Among the remaining ground sets in $\mathcal{S}$,
	no two ground sets have the same origin.
	Otherwise,
   the smaller ground set is either a singleton,
   which contradicts the bound $\ell+2$ on the number of singletons,
	or we have joined nodes to it at some step,
	contradicting our choice of the largest $C_i$ at step $i$.
	Hence, every  other ground set in
	$\mathcal{S}$ has grown \emph{maximally}.
	There are at most $\ell-1$ maximally grown ground sets,
	one per cluster,
	 excluding the cluster containing $B$. 
	Therefore,  $|\mathcal{S}| \leq 1 + 1 + (\ell+2)  + (\ell-1) = 2\ell+3$.
	Note that among all non-singleton ground sets in $\mathcal{S}$,
	 only $B$ does not grow during the process.
	 Thus,
    the total number of times that 
	a ground set in $\mathcal{S}$  has grown is
	$	\sum_{C^* \in \mathcal{S} } (|C^*|-1) - (k-1) $
	\begin{align*}
	=\sum_{C^* \in \mathcal{S} } |C^*| - 
		\sum_{C^* \in \mathcal{S} } 1 - (k-1) 
	\geq  k  \ell - (2\ell+3) - (k-1) =  (k-2)(\ell-1) - 4,
	\end{align*}
	which bounds the number of swaps performed by \ALG.
	The competitive ratio is then
	$\ALG / \OPT \geq ((k-2)(\ell-1) - 4) / 2$.
	%%
	\begin{comment}
	\ALG performs at least one swap  every time   a~ground set is constructed.
	Each expansion reduces the number of isolated nodes by $1$ or $2$.
	The decrease of two isolated nodes occurs when the largest component on  $I(x_i)$ is still singleton; this occurs at most $\ell$ times.
	We start with $k \cdot \ell - (k-1)$ isolated nodes (a~single $k-1$ ground set is revealed prior to $\{x_0, y_0\}$), and we end with isolated $\ell+1$ nodes.
	Hence, the total number of swaps of \ALG is at least $k\cdot \ell - k - 2 \ell$.
	The competitive ratio is then $\ALG / \OPT \geq (k\cdot \ell - k - 2\ell) / 2$.
	\end{comment}
	%%
\end{proof}

@@

	Finally, to bound the competitive ratio, we transform the above inequality in the following way: $|\sigma_E| \leq 2(\OPT(\sigma_E)+\OPT(M)) + 4\alpha \cdot \ell$.
	For succinctness, let $\xi := \OPT(\sigma_E) + \OPT(M)$.
	Combining the bounds on the cost of \TAlg and \OPT during each finished phase, the competitive ratio is
%
	\[
		\frac{\TAlg(\sigma)}{\OPT(\sigma)} \leq \frac{11\alpha \cdot \ell + |\sigma_E|}{\alpha/2 + \xi/2} \leq \frac{30\alpha\cdot\ell + 4\cdot \xi}{\alpha + \xi} \leq 30 \ell.
	\]
%	
	\medskip

	It remains to consider the last, unfinished phase.
	First, consider the case, where the unfinished phase is also the first one.
	Then, we cannot charge $\OPT$ due to the inability to partition the components.
	Instead, we use the fact that \TAlg and $\OPT$ started with the same initial partition.
	If the input finished before the first $\alpha$ external requests, then \TAlg is $1$-competitive.
	If at least $\alpha$ external requests were issued, then $\OPT$ either paid $\alpha$ for serving them remotely, or paid $\alpha$ for a~migration.
	Charging this cost to \OPT serves the purpose of charging $\alpha$ at the end of a finished phase, and thus we can repeat the analysis of a finished phase.
	Second, consider the case, where there are at least two phases, then we split the cost $\alpha$ charged in the penultimate phase into last two phases, and we repeat the analysis of a finished phase.
	This way, the competitive ratio increases at most twofold in comparison to a~finished phase, and the competitive ratio is $\TAlg(\sigma)/\OPT(\sigma) \leq 60\ell$.
\end{proof}
